### 2013-05-07 inst/examples/BUGS/allen.Rmd
added parametric bayesian myers and ricker for comparison

```{r sdp-pars, dependson="stateeq"}
sigma_g <- 0.05
sigma_m <- 0.0
z_g <- function() rlnorm(1, 0, sigma_g)
z_m <- function() 1+(2*runif(1, 0,  1)-1) * sigma_m
x_grid <- seq(0, 1.5 * K, length=50)
h_grid <- x_grid
profit <- function(x,h) pmin(x, h)
delta <- 0.01
OptTime <- 30  # stationarity with unstable models is tricky thing
reward <- 0
xT <- 0
Xo <- K # observations start from
x0 <- Xo # simulation under policy starts from
Tobs <- 40
```

@@

We define which parameters to keep track of, and set the initial values of parameters in the transformed space used by the MCMC.  We use logarithms to maintain strictly positive values of parameters where appropriate.  Because our priors on the noise parameters are inverse gamma distributed.

@@

```{r}
# a bit unfair to start with the correct values, but anyhow...
init_p = p
names(init_p) = c("r0", "K")
y <- obs$x[-1] 
N=length(y);
```

@@

@@

```{r}
# um, cleaner if we were just be using the long form, par_posterior
mcmc <- as.mcmc(jagsfit)
mcmcall <- mcmc[,-2]
who <- colnames(mcmcall)
who 
mcmcall <- cbind(mcmcall[,1],mcmcall[,2],mcmcall[,3],mcmcall[,4])
colnames(mcmcall) <- who
```

```{r}
ricker_pardist <- mcmcall
ricker_pardist[,2] = exp(ricker_pardist[,2]) # transform model parameters back first

bayes_coef <- apply(ricker_pardist,2,mean)
ricker_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[1]))
ricker_bayes_pars
```

@@

models <- data.frame(x=x_grid, GP=tgp_dat$y, True=true_means, 
                     MLE=est_means, Ricker=alt_means, 
                     Parametric.Bayes = par_bayes_means)

@@

```{r}
matrices_par_bayes <- parameter_uncertainty_SDP(f, p, x_grid, h_grid, pardist)
opt_par_bayes <- value_iteration(matrices_par_bayes, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
```

@@

```{r}
matrices_alt <- parameter_uncertainty_SDP(alt$f, ricker_bayes_pars[1,2], x_grid, h_grid, ricker_pardist)
opt_alt <- value_iteration(matrices_alt, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
```

@@

```{r}
OPT = data.frame(GP = opt_gp$D, True = opt_true$D, MLE = opt_estimated$D, Ricker = opt_alt$D, Parametric.Bayes = opt_par_bayes$D)
colorkey=cbPalette
names(colorkey) = names(OPT) 
```

@@

```{r Figure3}
ggplot(dt) + 
  geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=.1) +
  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1)))

>>>>

```{r sdp-pars, dependson="stateeq"}
sigma_g <- 0.05
sigma_m <- 0.0
z_g <- function() rlnorm(1, 0, sigma_g)
z_m <- function() 1+(2*runif(1, 0,  1)-1) * sigma_m
x_grid <- seq(0, 1.5 * K, length=50)
h_grid <- x_grid
profit <- function(x,h) pmin(x, h)
delta <- 0.01
OptTime <- 50  # stationarity with unstable models is tricky thing
reward <- 0
xT <- 0
Xo <- K # observations start from
x0 <- Xo # simulation under policy starts from
Tobs <- 40
```

@@

We define which parameters to keep track of, and set the initial values of parameters in the transformed space used by the MCMC.  We use logarithms to maintain strictly positive values of parameters where appropriate.

@@

```{r}
init_p = p
names(init_p) = c("r0", "K")
y <- obs$x[-1] 
N=length(y);
```

@@

```{r}
jags.data <- list("N","y")
n.chains = 1
n.iter = 400000
n.burnin = floor(10000)
n.thin = max(1, floor(n.chains * (n.iter - n.burnin)/1000))
```

@@

```{r}
ricker_pardist <- jags_matrix[! names(jags_matrix) == "deviance" ]
ricker_pardist[,"logr0"] = exp(ricker_pardist[,"logr0"]) # transform model parameters back first

posterior.mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

apply(ricker_pardist,2,mean)
bayes_coef <- apply(ricker_pardist,2, posterior.mode) # much better estimates
ricker_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[1]))
ricker_bayes_pars
```

## Write the external bugs file

```{r}
logr0_prior_p <- c(-6.0, 6.0)
logtheta_prior_p <- c(-6.0, 6.0)
logK_prior_p <- c(-6.0, 6.0)
stdQ_prior_p <- c(0.0001, 100)
stdR_prior_p <- c(0.0001, 100)

bugs.model <- 
paste(sprintf(
"model{
  logr0    ~ dunif(%s, %s)
  logtheta    ~ dunif(%s, %s)
  logK    ~ dunif(%s, %s)
  stdQ ~ dunif(%s, %s)
  stdR ~ dunif(%s, %s)", 
  logr0_prior_p[1], logr0_prior_p[2],
  logtheta_prior_p[1], logtheta_prior_p[2],
  logK_prior_p[1], logK_prior_p[2],
  stdQ_prior_p[1], stdQ_prior_p[2],
  stdR_prior_p[1], stdR_prior_p[2]), 
  "

  iQ <- 1 / (stdQ * stdQ);
  iR <- 1 / (stdR * stdR);

  r0 <- exp(logr0)
  theta <- exp(logtheta)
  K <- exp(logK)

  x[1] ~ dunif(0, 10)

  for(t in 1:(N-1)){
    mu[t] <- r0 * pow(abs(x[t]), theta) / (1 + pow(abs(x[t]), theta) / K)
    x[t+1] ~ dnorm(mu[t], iQ) 
  }

  for(t in 1:(N)){
    y[t] ~ dnorm(x[t], iR)
  }
}")

writeLines(bugs.model, "myers.bugs")
```

```{r}

## priors (untransformed variables)
logK_prior     <- function(x) dunif(x, logK_prior_p[1], logK_prior_p[2])
logr_prior     <- function(x) dunif(x, logr0_prior_p[1], logr0_prior_p[2])
logtheta_prior <- function(x) dunif(x, logtheta_prior_p[1], logtheta_prior_p[2])
stdQ_prior     <- function(x) dunif(x, stdQ_prior_p[1], stdQ_prior_p[2])
stdR_prior     <- function(x) dunif(x, stdR_prior_p[1], stdR_prior_p[2])

### Parametric Bayes based on myers model

Contains the

```{r}

init_p = c(1, 1, 1)
names(init_p) = c("r0", "theta", "K")
y <- obs$x[-1] 
N=length(y);
```

```{r}
jags.params=c("logr0", "logtheta", "logK", "stdQ", "stdR")
jags.inits <- function(){
  list("logr0"=log(init_p["r0"]), "logtheta" = log(init_p["theta"]),  
       "logK"=log(init_p["K"]), "stdQ"=sqrt(0.05), 
       "stdR"=sqrt(0.1),"x"=y,.RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}
```

```{r}
set.seed(12345)
time_jags <- system.time(       
  jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, 
                  n.chains=n.chains, n.iter=n.iter, n.thin=n.thin,
                  n.burnin=n.burnin, model.file="myers.bugs")
)         
time_jags <- unname(time_jags["elapsed"]);
```

Convergence diagnostics for parametric bayes

```{r}
jags_matrix <- as.data.frame(as.mcmc.bugs(jagsfit$BUGSoutput))
par_posteriors <- melt(cbind(index = 1:dim(jags_matrix)[1], jags_matrix), id = "index")

## Traces
ggplot(par_posteriors) + geom_line(aes(index, value)) +
  facet_wrap(~ variable, scale="free", ncol=1)

```{r}
## prior curve functions, must be in order
par_priors <- list( deviance = function(x) 0 * x, logK = logK_prior,
                    logr0 = logr_prior, logtheta = logtheta_prior, 
                    stdQ = stdQ_prior, stdR = stdR_prior)

## calculate points along the prior curve
par_prior_curves <- ddply(par_posteriors, "variable", function(dd){
    grid <- seq(min(dd$value), max(dd$value), length = 100) 
    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
})
```

```{r}
## posterior distributions
ggplot(par_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity", alpha=0.7) +
  geom_line(data=par_prior_curves, aes(x=value, y=density), col="red") + 
  facet_wrap(~ variable, scale="free", ncol=2)
```

```{r}
myers_pardist <- as.matrix(jags_matrix[2:6])
myers_pardist[,1] = exp(myers_pardist[,1]) # transform model parameters back first
myers_pardist[,2] = exp(myers_pardist[,2]) # transform model parameters back first
myers_pardist[,3] = exp(myers_pardist[,3]) # transform model parameters back first
colnames(myers_pardist) = c("K", "r0", "theta", "stdQ", "stdR")

apply(myers_pardist,2,mean)
bayes_coef <- apply(myers_pardist,2, posterior.mode) # much better estimates

myers_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[3], bayes_coef[1]))
```

@@

myers_means <- sapply(x_grid, Myer_harvest, 0, myers_bayes_pars)

models <- data.frame(x=x_grid, GP=tgp_dat$y, True=true_means, 
                     MLE=est_means, Ricker=alt_means, 
                     Parametric.Bayes = par_bayes_means,
                     Myers = myers_means)

@@

```{r}
allen_f <- function(x,h,p) unname(f(x,h,p[c(2, 1, 3)]))
matrices_par_bayes <- parameter_uncertainty_SDP(allen_f, x_grid, h_grid, pardist, 4)
opt_par_bayes <- value_iteration(matrices_par_bayes, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
```

@@

```{r}
ricker_f <- function(x, h, p) alt$f(x, h, unname(p[c(2, 1)]))
matrices_alt <- parameter_uncertainty_SDP(ricker_f, x_grid, h_grid, as.matrix(ricker_pardist), 3)
opt_alt <- value_iteration(matrices_alt, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
```

Bayesian Myers model

```{r}
myers_f <- function(x,h,p) Myer_harvest(x, h, p[c(2, 3, 1)])
matrices_myers <- parameter_uncertainty_SDP(myers_f, x_grid, h_grid, as.matrix(myers_pardist), 4)
myers_alt <- value_iteration(matrices_myers, x_grid, h_grid, OptTime=MaxT, xT, profit, delta=delta)
```

@@

```{r}
OPT = data.frame(GP = opt_gp$D, True = opt_true$D, MLE = opt_estimated$D, Ricker = opt_alt$D, Parametric.Bayes = opt_par_bayes$D, Myers = myers_alt$D)
colorkey=cbPalette
names(colorkey) = names(OPT) 
```

@@

```{r Figure3}
ggplot(dt) + 
  geom_line(aes(time, fishstock, group=interaction(reps,method), color=method), alpha=.1) +
  scale_colour_manual(values=colorkey, guide = guide_legend(override.aes = list(alpha = 1)))
```

```{r profits}
Profit <- dt[, sum(profit), by=c("reps", "method")]
Profit[, mean(V1), by="method"]
```

```{r totalprofits}
ggplot(Profit, aes(V1)) + geom_histogram() + 
  facet_wrap(~method, scales = "free_y") + guides(legend.position = "none")
```

### 2013-05-07 inst/examples/BUGS/bayesian-ricker.Rmd
added parametric bayesian myers and ricker for comparison

# Comparison of Nonparametric Bayesian Gaussian Process estimates to standard the Parametric Bayesian approach

### Parametric Bayes

We initiate the MCMC chain (`init_p`) using the true values of the parameters `p` from the simulation.  While impossible in real data, this gives the parametric Bayesian approach the best chance at succeeding.  `y` is the timeseries (recall `obs` has the $x_t$, $x_{t+1}$ pairs)

```{r}
# a bit unfair to start with the correct values, but anyhow...
init_p = p
names(init_p) = c("r0", "K")
y <- obs$x[-1] 
N=length(y);
```

@@

The actual model is defined in a `model.file` that contains an R function that is automatically translated into BUGS code by *R2WinBUGS*.  The file defines the priors and the model, as seen when read in here

```{r}
cat(readLines(con="ricker-UPrior.txt"), sep="\n")
```

@@

```{r}
# Uniform priors on standard deviation terms
jags.params=c("K","logr0", "stdQ", "stdR")
jags.inits <- function(){
  list("K"=init_p["K"],"logr0"=log(init_p["r0"]), "stdQ"=sqrt(0.05),"stdR"=sqrt(0.1),"x"=y,.RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}

set.seed(12345)

time_jags <- system.time(       
  jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, n.chains=n.chains, 
                  n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,model.file="ricker-UPrior.txt")
)         
time_jags <- unname(time_jags["elapsed"]);
```

@@

# Traces
ggplot(par_posteriors) + geom_line(aes(index, value)) + facet_wrap(~ variable, scale="free", ncol=1)

## priors (untransformed variables)
K_prior <- function(x) dunif(x, 0.01, 40)
logr_prior <- function(x) dunif(x, -6, 6)
stdQ_prior <- function(x) dunif(x, 0.001, 100)
stdR_prior <- function(x) dunif(x, 0.001, 100)

par_priors <- list(K = K_prior, deviance = function(x) 0 * x, logr0 = logr_prior, stdQ = stdQ_prior, stdR = stdR_prior)

par_prior_curves <- ddply(par_posteriors, "variable", function(dd){
    grid <- seq(min(dd$value), max(dd$value), length = 100) 
    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
})

# posterior distributions
ggplot(par_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity", alpha=0.7) +
  geom_line(data=par_prior_curves, aes(x=value, y=density), col="red") + 
  facet_wrap(~ variable, scale="free", ncol=2)

@@

```{r}
# um, cleaner if we were just be using the long form, par_posterior
mcmc <- as.mcmc(jagsfit)
mcmcall <- mcmc[,-2]
who <- colnames(mcmcall)
who 
mcmcall <- cbind(mcmcall[,1],mcmcall[,2],mcmcall[,3],mcmcall[,4])
colnames(mcmcall) <- who
```

```{r}
ricker_pardist <- mcmcall
ricker_pardist[,2] = exp(ricker_pardist[,2]) # transform model parameters back first

bayes_coef <- apply(ricker_pardist,2,mean)
ricker_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[1]))
ricker_bayes_pars
```

>>>>

```{r}
setwd("~/Documents/code/nonparametric-bayes/inst/examples/BUGS/")
opts_chunk$set(tidy = FALSE, warning = FALSE, message = FALSE, cache = FALSE, 
    comment = NA, fig.width = 6, fig.height = 4)
library(knitcitations)
library(ggplot2)  # plotting
opts_knit$set(upload.fun = socialR::flickr.url)
theme_set(theme_bw(base_size = 10))
theme_update(panel.background = element_rect(fill = "transparent", colour = NA), 
    plot.background = element_rect(fill = "transparent", colour = NA))
cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

Load necessary libraries,

```{r}
library(nonparametricbayes) # loads the rest as dependencies
```

### Model and parameters

Uses the model derived in `r citet("10.1080/10236190412331335373")`, of a Ricker-like growth curve with an allee effect, defined in the pdgControl package,

```{r}
f <- RickerAllee
p <- c(2, 8, 5)
K <- p[2]
allee <- p[3]
```

Various parameters defining noise dynamics, grid, and policy costs.

```{r}
sigma_g <- 0.05
sigma_m <- 0.0
z_g <- function() rlnorm(1, 0, sigma_g)
z_m <- function() 1+(2*runif(1, 0,  1)-1) * sigma_m
x_grid <- seq(0, 1.5 * K, length=50)
h_grid <- x_grid
profit <- function(x,h) pmin(x, h)
delta <- 0.01
OptTime <- 30  # stationarity with unstable models is tricky thing
reward <- 0
xT <- 0
Xo <- K # observations start from
x0 <- Xo # simulation under policy starts from
Tobs <- 40
```

### Sample Data

```{r}
  set.seed(1234)
  #harvest <- sort(rep(seq(0, .5, length=7), 5))
  x <- numeric(Tobs)
  x[1] <- 6
  nz <- 1
  for(t in 1:(Tobs-1))
    x[t+1] = z_g() * f(x[t], h=0, p=p)
  obs <- data.frame(x = c(rep(0,nz), 
                          pmax(rep(0,Tobs-1), x[1:(Tobs-1)])), 
                    y = c(rep(0,nz), 
                          x[2:Tobs]))
```

## Write the external bugs file

```{r}
logr0_prior_p <- c(-6.0, 6.0)
logtheta_prior_p <- c(-6.0, 6.0)
logK_prior_p <- c(-6.0, 6.0)
stdQ_prior_p <- c(0.0001, 100)
stdR_prior_p <- c(0.0001, 100)

bugs.model <- 
paste(sprintf(
"model{
  logr0    ~ dunif(%s, %s)
  logtheta    ~ dunif(%s, %s)
  logK    ~ dunif(%s, %s)
  stdQ ~ dunif(%s, %s)
  stdR ~ dunif(%s, %s)", 
  logr0_prior_p[1], logr0_prior_p[2],
  logtheta_prior_p[1], logtheta_prior_p[2],
  logK_prior_p[1], logK_prior_p[2],
  stdQ_prior_p[1], stdQ_prior_p[2],
  stdR_prior_p[1], stdR_prior_p[2]), 
  "

  iQ <- 1 / (stdQ * stdQ);
  iR <- 1 / (stdR * stdR);

  r0 <- exp(logr0)
  theta <- exp(logtheta)
  K <- exp(logK)

  x[1] ~ dunif(0, 10)

  for(t in 1:(N-1)){
    mu[t] <- r0 * pow(abs(x[t]), theta) / (1 + pow(abs(x[t]), theta) / K)
    x[t+1] ~ dnorm(mu[t], iQ) 
  }

  for(t in 1:(N)){
    y[t] ~ dnorm(x[t], iR)
  }
}")

writeLines(bugs.model, "meyers.bugs")
```

```{r}

## priors (untransformed variables)
logK_prior     <- function(x) dunif(x, logK_prior_p[1], logK_prior_p[2])
logr_prior     <- function(x) dunif(x, logr0_prior_p[1], logr0_prior_p[2])
logtheta_prior <- function(x) dunif(x, logtheta_prior_p[1], logtheta_prior_p[2])
stdQ_prior     <- function(x) dunif(x, stdQ_prior_p[1], stdQ_prior_p[2])
stdR_prior     <- function(x) dunif(x, stdR_prior_p[1], stdR_prior_p[2])

### Parametric Bayes based on Meyers model

Contains the

```{r}

init_p = c(1, 1, 1)
names(init_p) = c("r0", "theta", "K")
y <- obs$x[-1] 
N=length(y);
```

@@

@@

```{r}
jags.params=c("logr0", "logtheta", "logK", "stdQ", "stdR")
jags.inits <- function(){
  list("logr0"=log(init_p["r0"]), "logtheta" = log(init_p["theta"]),  
       "logK"=log(init_p["K"]), "stdQ"=sqrt(0.05), 
       "stdR"=sqrt(0.1),"x"=y,.RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}
```

```{r}
set.seed(12345)
time_jags <- system.time(       
  jagsfit <- jags(data=jags.data, inits=jags.inits, jags.params, 
                  n.chains=n.chains, n.iter=n.iter, n.thin=n.thin,
                  n.burnin=n.burnin, model.file="meyers.bugs")
)         
time_jags <- unname(time_jags["elapsed"]);
```

@@

## Traces
ggplot(par_posteriors) + geom_line(aes(index, value)) +
  facet_wrap(~ variable, scale="free", ncol=1)

@@

```{r}
## prior curve functions, must be in order
par_priors <- list( deviance = function(x) 0 * x, logK = logK_prior,
                    logr0 = logr_prior, logtheta = logtheta_prior, 
                    stdQ = stdQ_prior, stdR = stdR_prior)

## calculate points along the prior curve
par_prior_curves <- ddply(par_posteriors, "variable", function(dd){
    grid <- seq(min(dd$value), max(dd$value), length = 100) 
    data.frame(value = grid, density = par_priors[[dd$variable[1]]](grid))
})
```

```{r}
## posterior distributions
ggplot(par_posteriors, aes(value)) + 
  stat_density(geom="path", position="identity", alpha=0.7) +
  geom_line(data=par_prior_curves, aes(x=value, y=density), col="red") + 
  facet_wrap(~ variable, scale="free", ncol=2)
```

```{r}
meyers_pardist <- jags_matrix[2:6]
meyers_pardist[,2] = exp(meyers_pardist[,1]) # transform model parameters back first
meyers_pardist[,2] = exp(meyers_pardist[,2]) # transform model parameters back first
meyers_pardist[,2] = exp(meyers_pardist[,3]) # transform model parameters back first
colnames(meyers_pardist) = c("K", "r0", "theta", "stdQ", "stdR")

bayes_coef <- apply(meyers_pardist,2,mean)
meyers_bayes_pars <- unname(c(bayes_coef[2], bayes_coef[3], bayes_coef[1]))
```
